Gravitational Lens Recovery with \Glass: How to measure the mass profile and shape of a lens

### 2013-06-27 limits/ms.tex
Justin finished adding text/changes for now. Action items to be done listed in the conclusions.

This paper is organised as follows. In \secref{sec:theory}, we review the basic
lensing, stellar population synthesis, and Jeans mass modelling theory we will
need. We present some simple back-of-the-envelope calculations that illustrate
what type and quality of data we are likely to need to measure mass profiles
and shapes. In \secref{sec:glass}, we describe the \Glass\ lens inversion
algorithm and its stellar kinematic post-processing module. In
\secref{sec:mockdata}, we describe our mock N-body data and our ray tracing
algorithm used to generate the images. In \secref{sec:results}, we present our
results from applying \Glass\ to these mock data. Finally, in
\secref{sec:conclusions} we present our conclusions.

@@

Since we will later want to model the density distribution with a computer it
will be convenient to choose units that make the relevant quantities of order
unity.  We therefore measure lengths in light years, time in years, positions
in arcseconds, and choose $c=1$ and $4\pi G = N^2$, where $N^2 \equiv 206,265$
arcsec/rad. The mass unit is then $11.988\ \Msun$. It will also be useful to
define a proxy to the Hubble constant $\nu \equiv N^2 H_0$.

One way to understand the lens equation is via Fermat's principle. We can think of light as travelling only 
along extremum paths where lensed images occur \citep{1986ApJ...310..568B}. Such paths occur at the extrema of the photon {\it arrival time surface} $t(\vec\theta)$ that depends on the geometric path the photon takes and the general relativistic gravitational time dilation
due to a thin lens at redshift $z_L$. In its complete form, with all units
and cosmology included, we have \citep{1986ApJ...310..568B}:
%
\begin{eqnarray}
N^2ct(\vec\theta) & = & (1+z_L)\frac{D_{L}D_{S}}{D_{LS}}\frac12 |\vec\theta - \vec\beta|^2 \nonumber \\
& & - (1+z_L)\frac{4GD_{L}^2}{c^2}\int \Sigma(\vec\theta') \ln |\vec\theta-\vec\theta'| d^2\vec\theta'
\label{eqn:full_arrival_time}
\end{eqnarray}
%
where the factor of $D_L^2$ in the second term comes from the fact that $\Sigma$
has units of \Msun/lyr$^2$. We can clean this up by writing down a dimensionless time
delay:
%
\begin{equation}
\tau = \left[ (1+z_L) d_L\right]^{-1}\nu t
\label{tau}
\end{equation}
%
in terms of our previous definitions. If we
further define a dimensionless density:
%
\begin{equation}
\kappa_\infty = \frac{4\pi G}{c^2}\frac{c}{H_0}d_L\Sigma
              = \frac{d_L}{\nu}\Sigma
\end{equation}
%
and a lensing potential:
%
\begin{equation}
\psi(\vec\theta) = \frac1\pi \int \kappa_\infty(\vec\theta') \ln|\vec\theta - \vec\theta'| d^2\vec\theta'\
\label{lensing potential}
\end{equation}
%
we can express \eqnref{eqn:full_arrival_time} very compactly as:
%
\begin{equation}
\tau(\vec\theta) = \frac12 \xi |\vec\theta-\vec\beta|^2 - \psi(\vec\theta)
\label{arrival time}
\end{equation}
%
where $\xi \equiv D_{S}/D_{LS}$. We explicitly write $\kappa_\infty$ to remind ourselves
that there is no source distance factor involved. This will be useful later when we consider
multiple sources. Notice that the constraint $\nabla \tau = 0$ -- i.e. extrema in the arrival time surface -- recovers the lens equation (\eqnref{eqn:lens_equation}).

The only lensing observables from \eqnref{arrival time} are the image positions. However,  if the source object is
variable then the observed light curves of each image will be
identical but shifted in time due to the different photon path lengths.
The time difference between identical portions of the light curves is called
the time delay $\Delta t$ between images.  Using \eqnref{tau} the time delay
$\Delta t_{21}$ between two images $\vec\theta_1$ and $\vec\theta_2$, where $\vec\theta_2$
arrives after $\vec\theta_1$, is $\Delta t_{21} = \Delta \tau_{21}(1+z_L)d_L /
\nu$. If $\beta$ and $\kappa_\infty$ could be known exactly this would allow one
to precisely infer the angular diameter distance to the lens and thus -- via a cosmological model -- the Hubble constant. However, the source position is not
observable and the density distribution is highly degenerate.

@@

The trouble with solving the lens equation for the mass distribution is that the solutions are not unique. To see why this is, it is instructive to simplify \eqnref{eqn:lens_equation} for a circularly symmetric lens: 
%
\begin{equation}
\vec\beta = \vec\theta - \frac{\theta_E^2}{\theta^2}\vec\theta
\label{eqn:lens_equation_circ}
\end{equation}
%
where $\theta = |\vec\theta|$, $\theta_E^2 = \frac{D_{LS}}{D_S D_L} \frac{4 GM(<\theta)}{c^2}$ is the Einstein radius, and $M(<\theta)$ is the mass enclosed within the images.

For on-axis sources (\vec\beta\ = \vec0), the images split into a degenerate Einstein ring with $\theta = \theta_E$. Even in this situation of maximal information (since the source position $\vec\beta$ cannot typically be known), there is a trivial degeneracy between $M(<\theta_E)$ and the angular diameter distance. For fixed cosmology and in the most ideal situation, a single source splitting into images gives us only an enclosed mass. This `steepness' degeneracy (also known as the mass-sheet degeneracy) generalises to the non-circularly symmetric case \citep{1985ApJ...289L...1F,2000AJ....120.1654S}. Thus, to measure the mass profile of a lens, we require more information. This can come in the form of sources at different redshifts -- since each has its own Einstein radius, each constrains a different enclosed mass $M(<\theta_E')$. Or it can come in the form of additional constraints like time delay data between the images, stellar mass estimates from stellar populations synthesis modelling \citep{2008MNRAS.383..857F}, stellar kinematics \citep{2002MNRAS.337L...6T}, or similar. However, even with further information, higher order degeneracies remain. \citet{2008MNRAS.386..307L} showed that there is a monopole degeneracy that allows additional mass to be deposited in-between images without any optical effect, while \citep{2006ApJ...653..936S} find twisting/shape degeneracies that are sensitive to the time delays. Further degeneracies enter also if there is a strong {\it external} field acting on the lens -- for example from a larger group or cluster environment within which the lens sits \citep[e.g.][]{1996ApJ...468...17B}; this {\it shear term} can, in principle, however, be constrained by data \citep[e.g.][]{2002ApJ...578...25K}. In \secref{sec:results}, we will use \Glass\ to determine how well such degeneracies can be broken by data of ever-increasing quality to measure the mass profile and shape of a lens. We assume throughout that the shear is negligible.

@@

\begin{equation}
\sigma_p^2(R) = \frac{2}{I(R)}\int_R^\infty dr\nu(r)f(r)\frac{GM(r)}{r}
\label{eqn:sphericaljeans}
\end{equation}
where $\sigma_p$ is the projected velocity dispersion of the stars as a function of projected radius $R$; $I(R)$ is the surface density of the stars; $\nu(r)$ is the three dimensional stellar density; $f(r)$ is a function that depends only on the velocity anisotropy $\beta(r) = 1 - \sigma_t^2/\sigma_r^2$, where $\sigma_{t,r}$ are the tangential and radial velocity dispersions respectively:

\begin{equation}
f(r) = XXX;
\end{equation} 
$G$ is Newton's gravitational constant; and $M(r)$ is the mass profile that we would like to measure.

It is immediately clear from \eqnref{eqn:sphericaljeans} that, even assuming spherical symmetry, we have a degeneracy between the enclosed mass profile $M(r)$ and the velocity anisotropy $\beta(r)$. This can be understood intuitively since $\beta(r)$ measures the relative importance of radial versus circular orbits and is intrinsically difficult to constrain given only one component of the velocity vector for each star. Nonetheless, $\beta(r)$ can be constrained given sufficiently many stars, since radial Doppler velocities sample eccentric orbits as $r\rightarrow 0$ and tangential orbits as $r\rightarrow \infty$ (CITES). It can also be estimated if an independent measure of $M(r)$ is available -- for example coming from strong lensing. Finally, we may integrate out the effect of unknown $\beta(r)$ by performing a luminosity weighted average over all radii:

\begin{equation}
\overline{\sigma} = \frac{\int_0^\infty I(R) \sigma_p^2(R) 2\pi R dR}{\int_0^\infty I(R) 2\pi R dR}
\label{eqn:virialaverage}
\end{equation} 
This produces a single robust measure of the mass $\overline{\sigma}$ that is independent of $\beta(r)$, provided that the stellar system is in a dynamic quasi-equilibrium that obeys the virial theorem (CITES). In practice, we need not perform the above integral to infinity (this is impractical for real data since the surface brightness at large $R$ will eventually fall below the detection limit). Rather, we should see \eqnref{eqn:virialaverage} as a means to understand why fitting the projected velocity dispersion profile of stars gives us a single robust mass measure that is independent of $\beta(r)$ at some critical radius, typically $\sim$ the half light radius of the stellar distribution ($r_{1/2}$; CITES).

@@

\subsection{A post-processing module for stellar kinematics}\label{sec:glasskinematics} 
Similarly to the null-space constraint (\S\ref{sec:glassextraimages}), stellar kinematic constraints constitute a non-linear prior on the mass map and must be applied in post-processing. We sweep through the model ensemble performing an Abel deprojection to determine $M(r)$ from the projected surface density assuming spherical symmetry \citep[e.g.][]{2008gady.book.....B}:

\begin{equation} 
M(r) = XXX
\end{equation} 
This de-projection algorithm was tested on triaxial figures in \citet{2006ApJ...652L...5S}. They found that for triaxialities typical of our current cosmology, the method works extremely well unless the triaxial figure is projected directly along the line of sight such that we see the galaxy or galaxy cluster `down the barrel'. Such a situation is unlikely, but in any case avoidable since the resultant figure appears spherical in projection. This leads to the seemingly counter-intuitive result that the kinematic constraints -- that rely on the above de-projection -- are most secure for systems that do not appear spherical in projection (unless independent data can confirm the three dimensional shape is indeed very round).

@@

GOT TO HERE.

@@

The four mock galaxies we consider have been designed to test our recovery
procedure. They are two component models where the dark matter and stellar
profiles are allowed to be steep and shallow.  The enclosed mass in all cases
is fixed at $M(a_\star)= 1.8\e{10}$ at the scale radius $a_\star=2$ kpc. These
values were chosen to closely resemble B1115. We therefore place the galaxy at
a redshift of $z_L = 0.31$ for lensing.  Throughout, we assume a cosmology
where $H_0^{-1}=13.7$ Gyr, $\Omega_M=0.28$, and $\Omega_\Lambda=0.72$. The
critical density is $\kappa_\mathrm{crit}\sim 1.8\e{9}$\Msun/kpc$^2$.

The galaxies were
first generated as three dimensional particle distributions with the
tool DEHNEN-TOOL. Each component follows the profile
\begin{equation}
\rho(r) = \frac{M}{4\pi a^3}(3-\gamma){(r/a)^{-\gamma}(1 + r/a)^{\gamma-4}}
\label{Dehnen profile}
\end{equation}
where $\gamma$ is the profile index and $a$ is the scale radius.  In the
case $\gamma=1$ this is the Hernquist profile.  The four combinations of
profile indices are shown in \tabref{mock galaxy params}.  Note that two of
the galaxies have $\gamma_\mathrm{DM}=1$. While we know from simulations that a
more accurate approximation would be an NFW or Einasto profile we are not
probing the outer regions of the halo where the difference would be most
extreme.  The Hernquist profile is easier to generate as it has finite mass.

\begin{table}
\begin{tabular}{cllllllll}
Galaxy & $\gamma_\star$ & $M_\star$ & $\gamma_\mathrm{DM}$ & $M_\mathrm{DM}$ & $\Rmap$ & Notes\\
\hline
AA & 1 & 4 & 0.05 & $11^{2.95}$ & 50 kpc & \\
AC & 1 & 4 & 1 & $11^2$ & 50 kpc & \\
BB & 1.5 & $2^{1.5}$ & 0.16 & $11^{2.84}$ & 50 kpc & \\
BC & 1.5 & $2^{1.5}$ & 1 & $11^2$ & 10 kpc & 
\end{tabular}
\caption{Profile parameters for the four mock galaxies. The resulting profiles only roughly follow
\eqnref{Dehnen profile} because the galaxies are triaxial. Masses are in units of $1.8\e{10}\Msun$. The scale lengths for
all lenses are $(a_\star,a_\mathrm{DM})=(2,20)$ kpc and by definition
$M_\star(a_\star) = M_\mathrm{DM}(a_\star) = 1$. $\Rmap$ is the 2d projected radius used to generate the lens configurations.}
\label{mock galaxy params}
\end{table}

\begin{figure}
\plotthree{MockGalProfile-a.pdf} {MockGalProfile-b.pdf} {MockGalProfile-c.pdf}
%\plotone{MockGalProfile-a.pdf} \plotone{MockGalProfile-b.pdf} \plotone{MockGalProfile-c.pdf}
\caption{
\textbf{(Left)} 
Three-dimensional density of the galaxies showing the stellar, dark matter,
and total densities with dotted, dashed, and solid lines, respectively. The
grey lines show the analytic forms from \eqnref{Dehnen profile}. 
\textbf{(Middle)} 
The radially averaged two-dimensional projected density.
The critical density for lensing at $z_L=0.31$ is $\kappa_\mathrm{crit}\sim 1.8\e{9}$\Msun/kpc$^2$.
\textbf{(Right)}
The enclosed projected mass.
}
\label{mock galaxies}
\end{figure}

In \figref{mock galaxies} we show the 3D radial density, 
the 2D projected density, and the 2D enclosed mass for each
galaxy. Solid lines represent the total mass, dashed lines the dark 
matter, and dotted lines the stellar mass. The projected plot has
been normalized to the critical density.

\subsection{Lens configurations} %--------------------------------------------------------------

For each of the four galaxies, we used the raytracing feature of \Glass\
described in \secref{Raytracing} to construct 12 lensing scenarios:

\begin{enumerate}
\item One double and one extended double.
\item One quad and one extended quad.
\item Two 2-source quads with varying redshift contrast.
\end{enumerate}

The extended configurations use extra sources to generate extra images that
simulate an arc. \figref{config ex.} shows the configuration for the extended
quad.

Each of these configurations were modelled with and without time delays and with
and without a central image. We assumed for all our tests that the lensing mass
was radially symmetric. By construction, this is known to be true and this is
most often the case with real galaxies, unless there is an obvious observed
asymmetry. The central pixel was refined into a further nine pixels to capture
any steep rise in the profile. Two of the four mock galaxies have a steeply rising
inner profile.

\section{Results}\label{sec:results}

\figref{reconstruction} shows a typical reconstruction of a
lens. The far left plot shows the ensemble average arrival time surface with
image marked as circles and the inferred source position as a diamond. The
center plot shows the radial density profile. The error bars cover the full
range of models generated by \Glass. The true density profile from the mock
data is also plotted for comparison. The vertical lines mark the radial
position of the images. The final plot on the right is of the enclosed mass. As
expected the error bars are smallest in the region of the images where the most
information about the lens is present. The dip in the profile at the end of the
profile is due to the cut off in mass in the lensing map. This is of little
importance, though, as there is no lensing information there.

\begin{figure}
\plotthree{BCQuadR1a_TmS-a.pdf}{BCQuadR1a_TmS-b.pdf}{BCQuadR1a_TmS-c.pdf}
\caption{A typical reconstructed lens. Here we present a single quad lens from 
the BC mock galaxy.
\textbf{(Left)}
The arrival time surface. 
\textbf{(Middle)}
The surface density. The magenta curve represents the dark matter component,
the yellow curve the stellar component, and the blue curve is the sum of the two.
The black curve comes from the original mass model used to create the lens.
The green vertical lines mark the radial positions of the images. The higher
resolution feature of \Glass\ has been used on the central pixel allowing the
steep profile to be captured.
\textbf{(Right)}
The cumulative mass.}
\label{reconstruction}
\end{figure}

The main results from modelling these different configuration are shown black in
\figref{main results}. Each subplot corresponds to a different galaxy and
the vertical axis shows the range of quality.  The quality of a
model $\M_i$ by comparing the recovered density profile $\rho_\M(r)$ against
the profile of the mock galaxy $\rho_G(r)$.  In particular, the plots show the
RMS quantity
%
\begin{equation}
  \chi_i = 100 \times \sqrt{\frac{\int_r (\rho_{\M_i}(r) - \rho_G(r))^2}{\int_r \rho_G(r)^2}}
  %\chi_i = \frac{\sum_r (\rho(r)_{\M_i} - \rho(r)_G)^2}{\sum_r \rho(r)_G^2}
\end{equation}
%

\begin{figure}
\plottwo{AAchi2_profile-full.pdf}{BBchi2_profile-full.pdf}

\plottwo{ACchi2_profile-full.pdf}{BCchi2_profile-full.pdf}
\caption{The main results showing the quality of model recovery. Each panel corresponds to 
the named mock galaxy, whose parameters are listed in \tabref{mock galaxy params}. Within
each panel are six groups of results for each of six lens morphologies. Each morphology
considered the presence of time delays and a central image. The black markers are for tests
that did not include the stellar mass as a lower bound constraint, while the red markers
indicate where the stellar mass has been given.}
\label{main results}
\end{figure}

\begin{figure}
\plottwo{AAchi2_profile.pdf}{BBchi2_profile.pdf}

\plottwo{ACchi2_profile.pdf}{BCchi2_profile.pdf}
\caption{Profile $\chi$ with $1\sigma$ error bars.}
\label{main results}
\end{figure}

\begin{figure}
\plottwo{AAchi2-full.pdf}{BBchi2-full.pdf}

\plottwo{ACchi2-full.pdf}{BCchi2-full.pdf}
\caption{Pixel-wise $\chi$.}
\label{main results}
\end{figure}

>>>>

This paper is organised as follows. In \secref{sec:theory}, we review the basic
lensing, stellar population synthesis, and Jeans mass modelling theory we will
need. We present some simple back-of-the-envelope calculations that illustrate
what type and quality of data we are likely to need to measure mass profiles
and shapes. In \secref{sec:glass}, we describe the \Glass\ lens inversion
algorithm and its stellar kinematic post-processing module. In
\secref{sec:mockdata}, we describe our mock N-body data. In \secref{sec:results}, we present our
results from applying \Glass\ to these mock data. Finally, in
\secref{sec:conclusions} we present our conclusions.

@@

Since we will later want to model the density distribution with a computer it
will be convenient to choose units that make the relevant quantities of order
unity.  We therefore measure lengths in light years, time in years, positions
in arcseconds, and choose $c=1$ and $4\pi G = N^2$, where $N^2 \equiv 206,265$
arcsec/rad. The mass unit is then $11.988\ \Msun$. It will also be useful to
define a proxy to the Hubble constant $\zeta \equiv N^2 H_0$.

One way to understand the lens equation is via Fermat's principle. We can think of light as travelling only 
along extremum paths where lensed images occur \citep{1986ApJ...310..568B}. Such paths occur at the extrema of the photon {\it arrival time surface} $t(\vec\theta)$ that depends on the geometric path the photon takes and the general relativistic gravitational time dilation
due to a thin lens at redshift $z_L$. In its complete form, with all units
and cosmology included, we have \citep{1986ApJ...310..568B}:
%
\begin{eqnarray}
N^2ct(\vec\theta) & = & (1+z_L)\frac{D_{L}D_{S}}{D_{LS}}\frac12 |\vec\theta - \vec\beta|^2 \nonumber \\
& & - (1+z_L)\frac{4GD_{L}^2}{c^2}\int \Sigma(\vec\theta') \ln |\vec\theta-\vec\theta'| d^2\vec\theta'
\label{eqn:full_arrival_time}
\end{eqnarray}
%
where the factor of $D_L^2$ in the second term comes from the fact that $\Sigma$
has units of \Msun/lyr$^2$. We can clean this up by writing down a dimensionless time
delay:
%
\begin{equation}
\tau = \left[ (1+z_L) d_L\right]^{-1}\zeta t
\label{tau}
\end{equation}
%
in terms of our previous definitions. If we
further define a dimensionless density:
%
\begin{equation}
\kappa_\infty = \frac{4\pi G}{c^2}\frac{c}{H_0}d_L\Sigma
              = \frac{d_L}{\zeta}\Sigma
\end{equation}
%
and a lensing potential:
%
\begin{equation}
\psi(\vec\theta) = \frac1\pi \int \kappa_\infty(\vec\theta') \ln|\vec\theta - \vec\theta'| d^2\vec\theta'\
\label{lensing potential}
\end{equation}
%
we can express \eqnref{eqn:full_arrival_time} very compactly as:
%
\begin{equation}
\tau(\vec\theta) = \frac12 \xi |\vec\theta-\vec\beta|^2 - \psi(\vec\theta)
\label{arrival time}
\end{equation}
%
where $\xi \equiv D_{S}/D_{LS}$. We explicitly write $\kappa_\infty$ to remind ourselves
that there is no source distance factor involved. This will be useful later when we consider
multiple sources. Notice that the constraint $\nabla \tau = 0$ -- i.e. extrema in the arrival time surface -- recovers the lens equation (\eqnref{eqn:lens_equation}).

The only lensing observables from \eqnref{arrival time} are the image positions. However,  if the source object is
variable then the observed light curves of each image will be
identical but shifted in time due to the different photon path lengths.
The time difference between identical portions of the light curves is called
the time delay $\Delta t$ between images.  Using \eqnref{tau} the time delay
$\Delta t_{21}$ between two images $\vec\theta_1$ and $\vec\theta_2$, where $\vec\theta_2$
arrives after $\vec\theta_1$, is $\Delta t_{21} = \Delta \tau_{21}(1+z_L)d_L /
\zeta$. If $\beta$ and $\kappa_\infty$ could be known exactly this would allow one
to precisely infer the angular diameter distance to the lens and thus -- via a cosmological model -- the Hubble constant. However, the source position is not
observable and the density distribution is highly degenerate.

@@

The trouble with solving the lens equation for the mass distribution is that the solutions are not unique. To see why, it is instructive to simplify \eqnref{eqn:lens_equation} for a circularly symmetric lens: 
%
\begin{equation}
\vec\beta = \vec\theta - \frac{\theta_E^2}{\theta^2}\vec\theta
\label{eqn:lens_equation_circ}
\end{equation}
%
where $\theta = |\vec\theta|$, $\theta_E^2 = \frac{D_{LS}}{D_S D_L} \frac{4 GM(<\theta)}{c^2}$ is the Einstein radius, and $M(<\theta)$ is the mass enclosed within the images.

For on-axis sources (\vec\beta\ = \vec0), the images split into a degenerate Einstein ring with $\theta = \theta_E$. Even in this situation of maximal information (since the source position $\vec\beta$ cannot typically be known), there is a trivial degeneracy between $M(<\theta_E)$ and the angular diameter distance. For fixed cosmology and in the most ideal situation, a single source splitting into images gives us only an enclosed mass. This `steepness' degeneracy (also known as the mass-sheet degeneracy) generalises to the non-circularly symmetric case \citep{1985ApJ...289L...1F,2000AJ....120.1654S}. Thus, to measure the mass profile of a lens, we require more information. This can come in the form of sources at different redshifts -- since each has its own Einstein radius, each constrains a different enclosed mass $M(<\theta_E')$. Or it can come in the form of additional constraints like time delay data between the images, stellar mass estimates from stellar populations synthesis modelling \citep{2008MNRAS.383..857F}, stellar kinematics \citep{2002MNRAS.337L...6T}, or similar. However, even with further information, higher order degeneracies remain. \citet{2008MNRAS.386..307L} showed that there is a monopole degeneracy that allows additional mass to be deposited in-between images without any optical effect, while \citep{2006ApJ...653..936S} find twisting/shape degeneracies that are sensitive to the time delays. Further degeneracies enter also if there is a strong {\it external} field acting on the lens -- for example from a larger group or cluster environment within which the lens sits \citep[e.g.][]{1996ApJ...468...17B}. However, this {\it shear term} can be constrained by data \citep[e.g.][]{2002ApJ...578...25K}. In \secref{sec:results}, we will use \Glass\ to determine how well such degeneracies can be broken by data of ever-increasing quality to measure the mass profile and shape of a lens. We assume throughout that the shear is negligible.

@@

\begin{equation}
\sigma_p^2(R) = \frac{2}{I(R)}\int_R^\infty dr \left(1-\beta \frac{R^2}{r^2}\right) \frac{\nu \sigma_r^2 r}{\sqrt{r^2 - R^2}};
\label{eqn:sphericaljeans}
\end{equation}
\begin{equation} 
\sigma_r^2(r) = \frac{r^{-2\beta}}{\nu}\int_r^\infty r'^{2\beta} \nu \frac{GM(r')}{r'^2}dr'
\end{equation} 
where $\sigma_p$ is the projected velocity dispersion of the stars as a function of projected radius $R$; $I(R)$ is the surface density of the stars; $\nu(r)$ is the three dimensional stellar density; $\sigma_{r,t}(r)$ are the radial and tangential velocity dispersions, respectively; $\beta(r) = 1 - \sigma_t^2/\sigma_r^2 = \mathrm{const.}$ is the velocity anisotropy (here assumed to be constant); $G$ is Newton's gravitational constant; and $M(r)$ is the mass profile that we would like to measure.

It is immediately clear from \eqnref{eqn:sphericaljeans} that, even assuming spherical symmetry, we have a degeneracy between the enclosed mass profile $M(r)$ and the velocity anisotropy $\beta(r)$. This can be understood intuitively since $\beta(r)$ measures the relative importance of radial versus circular orbits and is intrinsically difficult to constrain given only one component of the velocity vector for each star. Nonetheless, $\beta(r)$ can be constrained given sufficiently many stars, since radial Doppler velocities sample eccentric orbits as $r\rightarrow 0$ and tangential orbits as $r\rightarrow \infty$ \citep[e.g.][]{2002MNRAS.330..778W}. It can also be estimated if an independent measure of $M(r)$ is available -- for example coming from strong lensing. Finally, we may integrate out the effect of unknown $\beta(r)$ by performing a luminosity weighted average over all radii:

\begin{equation}
\overline{\sigma} = \frac{\int_0^{a_p} I(R) \sigma_p^2(R) 2\pi R dR}{\int_0^{a_p} I(R) 2\pi R dR}
\label{eqn:virialaverage}
\end{equation} 
where $a_p$ is the `aperture' over which the integral is performed. Ideally we should take $a_p \rightarrow \infty$, however this is not practical for real data where the surface brightness of the stars falls off rapidly with $a_p$. In practice, $a_p > 2r_{1/2}$, where $r_{1/2}$ is the half light radius of the stars, appears to be sufficient (see \S\ref{sec:results}). The above average produces a single robust measure of the mass $\overline{\sigma}$ that is independent of $\beta(r)$, provided that the stellar system is in a dynamic quasi-equilibrium that obeys the virial theorem \citep{2009ApJ...704.1274W,2010MNRAS.406.1220W,2012ApJ...754L..39A}. (Such an averaging procedure was first used in a combined lensing/stellar kinematic analysis in \citealt{2002MNRAS.337L...6T}.)

@@

{\bf JR. Jonathan to fill in this section.}

\subsection{A post-processing module for stellar kinematics}\label{sec:glasskinematics} 
Similarly to the null-space constraint (\S\ref{sec:glassextraimages}), stellar kinematic constraints constitute a non-linear prior on the mass map and must be applied in post-processing. We sweep through the model ensemble performing an Abel deprojection to determine $M(r)$ from the projected surface density assuming spherical symmetry \citep[e.g.][]{2008gady.book.....B,2008MNRAS.390.1647B}:

\begin{eqnarray} 
M(r) & = & \frac{\zeta}{d_L} \left[2\pi\int_0^r R\kappa_\infty(R)dR - \right. \nonumber \\
& & \left. 4\int_r^\infty R\kappa_\infty(R)f\left(\frac{R}{r}\right)dR\right]
\end{eqnarray} 
This de-projection algorithm was tested on triaxial figures in \citet{2006ApJ...652L...5S}. They found that for triaxialities typical of our current cosmology, the method works extremely well unless the triaxial figure is projected directly along the line of sight such that we see the galaxy or galaxy cluster `down the barrel'. Such a situation is unlikely, but in any case avoidable since the resultant figure appears spherical in projection. This leads to the seemingly counter-intuitive result that the kinematic constraints -- that rely on the above de-projection -- are most secure for systems that do not appear spherical in projection (unless independent data can confirm the three dimensional shape is indeed very round).

@@

@@

We generate four two-component mock galaxies, where the dark matter and stellar
profiles are allowed to be both steep and shallow.  The enclosed mass in all cases
is fixed at $M(a_\star)= 1.8\e{10}$\,M$_\odot$ at the scale radius $a_\star=2$\,kpc. These
values were chosen to closely resemble the lensing galaxy B1115 \citep{1980Natur.285..641W}. We place the galaxy at
a redshift of $z_L = 0.31$ for lensing.  Throughout, we assume a cosmology
where $H_0^{-1}=13.7$ Gyr, $\Omega_M=0.28$, and $\Omega_\Lambda=0.72$. The
critical density is $\kappa_\mathrm{crit}\sim 1.8\e{9}$\Msun/kpc$^2$.

The galaxies were generated as three dimensional particle distributions as in \citet{2009MNRAS.395.1079D}. Each component follows the profile:
\begin{equation}
\rho(m) = \frac{M}{4\pi a^3}(3-\gamma){(m/a)^{-\gamma}(1 + m/a)^{\gamma-4}}
\label{Dehnen profile}
\end{equation}
where $m^2 = (x/a)^2 + (y/b)^2 + (z/c)^2$, with $[a,b,c] = ???$; $\gamma$ is the central density profile index; and $a$ is the scale radius. In the
case $\gamma=1$ (and in the limit of spherical symmetry), this is the Hernquist profile \citep{1990ApJ...356..359H}.  The four combinations of
profile indices are shown in \tabref{mock galaxy params}. {\bf JR. Need to add explanation of triaxial parameters here. Is the above density equation correct in fact?}

\begin{table}
\begin{tabular}{cllllllll}
Galaxy & $\gamma_\star$ & $M_\star$ & $\gamma_\mathrm{DM}$ & $M_\mathrm{DM}$ & $\Rmap$ & Notes\\
\hline
AA & 1 & 4 & 0.05 & $11^{2.95}$ & 50 kpc & DM; shallow\\
AC & 1 & 4 & 1 & $11^2$ & 50 kpc & DM; cuspy\\
BB & 1.5 & $2^{1.5}$ & 0.16 & $11^{2.84}$ & 50 kpc &  Star; shallow\\
BC & 1.5 & $2^{1.5}$ & 1 & $11^2$ & 10 kpc & Star; cuspy
\end{tabular}
\caption{Profile parameters for the four mock galaxies. Masses are in units of $1.8\e{10}\Msun$. The scale lengths for
all lenses are $(a_\star,a_\mathrm{DM})=(2,20)$\,kpc. $\Rmap$ is the 2D projected radius used to generate the lens configurations. In the notes column we describe the mock as DM or star dominated with a shallow or cuspy central dark matter density profile.}
\label{mock galaxy params}
\end{table}

\begin{figure*}
\includegraphics[width=0.33\textwidth]{MockGalProfile-a.pdf} 
\includegraphics[width=0.33\textwidth]{MockGalProfile-b.pdf} 
\includegraphics[width=0.33\textwidth]{MockGalProfile-c.pdf}
\caption{{\bf JR. Cut the analytic profiles from this Figure. All fonts and labels should be larger. Left panel should start at 0.7kpc.}
\textbf{Left:} 
Density averaged in spherical shells of the four galaxy mocks: AA, AC, BB and BC, showing the stellar (dotted), dark matter (dashed),
and total (solid) densities, respectively. Notice that the stars in models BB and BC contribute significantly to the central potential. 
\textbf{Middle:} 
The radially averaged two-dimensional projected density.
The critical density for lensing at $z_L=0.31$, $\kappa_\mathrm{crit}\sim 1.8\e{9}$\Msun/kpc$^2$, is marked by the horizontal line. 
\textbf{Right:}
The enclosed projected mass.
}
\label{mock galaxies}
\end{figure*}

In \figref{mock galaxies}, we show the 3D radial density, 
the 2D projected density, and the 2D enclosed mass for each
galaxy.

\subsection{Lens configurations}\label{sec:lensconfig} %--------------------------------------------------------------

For each of the four galaxies, we used the raytracing feature of \Glass\
described in \secref{Raytracing} to construct 12 lensing scenarios {\bf JR. We have 12 lensing scenarios, but only 6 panels in Figure 2. Is this number correct?}:

\begin{enumerate}
\item one double and one extended double;
\item one quad and one extended quad;
\item two 2-source quads with varying redshift contrast.
\end{enumerate}
The `extended' configurations use multiple point sources at the same redshift to simulate an extended source (that will produce an arc-like image). \figref{arrival surfaces} shows the configuration for the extended quad, along with the other cases.

\begin{figure*}
\includegraphics[width=0.85\textwidth]{BCarrival_surfaces}
\caption{The lens configurations for the six test cases using the BC mock galaxy. Here,
the central image is shown, although not all tests include it. The central image belongs
to only one set of images to avoid overconstraining the models. Similar colors group
images that share a common source. In the case of the extended image systems the source
is at the same redshift, while the redshift is varied between the systems in the 
ZConstrast cases. Grey circles are a visual aid to help determine radial separation
between images. The axes are measure in arcseconds.}
\label{arrival surfaces}
\end{figure*}

Each of these configurations were modelled with and without time delays and with
and without a central image. (The central image is typically highly demagnified. For galaxy lenses it is very difficult to find since it lies along the sight line to the bright lensing galaxy; in clusters, however, such images have been seen (CITES)). We assumed for all our tests that the lensing mass
was radially symmetric ({\bf JR. Need to better-define this symmetry here.}. For our mock data, this is known to be true; it is
most often the case with real galaxies, unless there is an obvious observed
asymmetry. The central pixel was refined into a further nine pixels to capture
any steep rise in the profile. Two of the four mock galaxies have a steeply rising
inner profile. {\bf JR. What happens if we don't assume this symmetry? We ought to add an appendix showing this for one example case.}

\section{Results}\label{sec:results}
\figref{reconstruction} shows a typical reconstruction of a
lens. The far left plot shows the ensemble average arrival time surface with
image marked as circles and the inferred source position as a diamond. The
center plot shows the radial density profile. The error bars cover the full
range of models generated by \Glass {\bf JR. Why do this? Surely 1 or 2 sigma errors make more sense? I've deleted the ``full ensemble" results and kept the $1\sigma$ ones instead.}. The true density profile from the mock
data is also plotted for comparison. The vertical lines mark the radial
position of the images. The final plot on the right is of the enclosed mass. As
expected, the error bars are smallest in the region of the images where the most
information about the lens is present. The dip in the profile at the end of the
profile is due to the cut off in mass in the lensing map. This is of little
importance, though, as there is no lensing information there.

\begin{figure*}
\includegraphics[width=0.33\textwidth]{BCQuadR1a_TmS-a.pdf}
\includegraphics[width=0.33\textwidth]{BCQuadR1a_TmS-b.pdf}
\includegraphics[width=0.33\textwidth]{BCQuadR1a_TmS-c.pdf}
\caption{
{\bf JR. Need to homogenise plot dimensions here. Larger fonts as in Figure 1 are also required.} 
A typical reconstructed lens. Here we present results for a single quad image lensed by the BC mock galaxy.
\textbf{Left:}
The arrival time surface. 
\textbf{Middle:}
The surface density. The magenta curve represents the dark matter component,
the yellow curve the stellar component, and the blue curve is the sum of the two.
The black curve comes from the original mass model used to create the lens.
The green vertical lines mark the radial positions of the images. The higher
resolution feature of \Glass\ has been used on the central pixel allowing the
steep profile to be captured.
\textbf{Right:}
The cumulative mass.}
\label{reconstruction}
\end{figure*}
